ppseq: An R Package for Sequential Predictive Probability Monitoring

### 2022-10-03 inst/ppseq-package-paper/zabor-hobbs-kane.Rmd
first paper revision draft

The case study focuses on a re-design of a phase II study of atezolizumab in metastatic urothelial carcinoma patients (mUC) using sequential predictive probability monitoring. Atezolizumab is a programmed death-ligand 1 (PD-L1) blocking monoclonal antibody that was given accelerated approval by the U.S. Food and Drug Administration in May 2016 for the treatment of patients with locally advanced or metastatic urothelial carcinoma who had disease progression following platinum-containing chemotherapy. The approval was based on the results of a single-arm phase II study in 310 patients [@Rosenberg2016]. The phase II study used a hierarchical fixed-sequence testing procedure to test increasingly broad subgroups of patients based on PD-L1 status, and found overall response rates of 26\% (95\% CI: 18-36), 18\% (95\% CI: 13-24), and 15\% (95\% CI 11-19) in patients with ≥5\% PD-L1-positive immune cells (IC2/3 subgroup), in patients with ≥1\% PD-L1-positive immune cells (IC1/2/3 subgroup), and in all patients, respectively [@Rosenberg2016]. All three rates exceeded the historical control rate of 10\%. Then, in March 2021, the approval in this indication was voluntarily withdrawn by the sponsor following negative results from a randomized phase III study [@Powles2018]. In the phase III study, 931 patients were randomly assigned to receive atezolizumab or chemotherapy in a 1:1 ratio, and the same hierarchical fixed-sequence testing procedure as in the phase II study was used. The phase III study found that overall survival did not differ significantly between the atezolizumab and chemotherapy groups of the IC2/3 subgroup (median survival 11.1 months [95\% CI: 8.6-15.5] versus 10.6 months [95\% CI: 8.4-12.2]), so no further testing was conducted for the primary endpoint [@Powles2018]. Further analyses revealed that while the response rates to atezolizumab were comparable to those seen in the phase II study, the response rates to chemotherapy were much higher than the historical control rate of 10\%. The overall response rates to chemotherapy were 21.6\% (95\% CI: 14.5-30.2), 14.7\% (95% CI: 10.9-19.2), and 13.4\% (95\% CI: 10.5-16.9) for the IC2/3 subgroup, IC1/2/3 subgroup, and all patients, respectively. The overall response rates to atezolizumab were 23\% (95\% CI: 15.6-31.9), 14.1\% (95\% CI: 10.4-18.5), and 13.4\% (95\% CI: 10.5-16.9) for the IC2/3 subgroup, IC1/2/3 subgroup, and all patients, respectively. These results indicate that PD-L1 status is a prognostic biomarker for both standard of care chemotherapies that comprised the control arm as well as atezolizumab in this patient population.

@@

```{r message = FALSE, fig.height = 9, fig.width = 6, include = knitr::is_latex_output(), eval = knitr::is_latex_output(), fig.cap = "Plot of decision rules made with ggplot. The color indicates whether the trial should stop or proceed for a given number of responses at each interim analysis."}
plot(two_sample_decision_tbl, plotly = FALSE)
```

```{r message = FALSE, fig.height = 9, fig.width = 6, include=knitr::is_html_output(), eval=knitr::is_html_output(), fig.cap = "Plot of decision rules made with plotly. The color indicates whether the trial should stop or proceed for a given number of responses at each interim analysis."}
plot(two_sample_decision_tbl)
```

>>>>

The case study focuses on a re-design of a phase II study of atezolizumab in metastatic urothelial carcinoma patients (mUC) using sequential predictive probability monitoring. Atezolizumab is a programmed death-ligand 1 (PD-L1) blocking monoclonal antibody that was given accelerated approval by the U.S. Food and Drug Administration in May 2016 for the treatment of patients with locally advanced or metastatic urothelial carcinoma who had disease progression following platinum-containing chemotherapy. The approval was based on the results of a single-arm phase II study in 310 patients [@Rosenberg2016]. The phase II study used a hierarchical fixed-sequence testing procedure to test increasingly broad subgroups of patients based on PD-L1 status, and found overall response rates of 26\% (95\% CI: 18-36), 18\% (95\% CI: 13-24), and 15\% (95\% CI 11-19) in patients with $\geq5\%$ PD-L1-positive immune cells (IC2/3 subgroup), in patients with $\geq1\%$ PD-L1-positive immune cells (IC1/2/3 subgroup), and in all patients, respectively [@Rosenberg2016]. All three rates exceeded the historical control rate of 10\%. Then, in March 2021, the approval in this indication was voluntarily withdrawn by the sponsor following negative results from a randomized phase III study [@Powles2018]. In the phase III study, 931 patients were randomly assigned to receive atezolizumab or chemotherapy in a 1:1 ratio, and the same hierarchical fixed-sequence testing procedure as in the phase II study was used. The phase III study found that overall survival did not differ significantly between the atezolizumab and chemotherapy groups of the IC2/3 subgroup (median survival 11.1 months [95\% CI: 8.6-15.5] versus 10.6 months [95\% CI: 8.4-12.2]), so no further testing was conducted for the primary endpoint [@Powles2018]. Further analyses revealed that while the response rates to atezolizumab were comparable to those seen in the phase II study, the response rates to chemotherapy were much higher than the historical control rate of 10\%. The overall response rates to chemotherapy were 21.6\% (95\% CI: 14.5-30.2), 14.7\% (95% CI: 10.9-19.2), and 13.4\% (95\% CI: 10.5-16.9) for the IC2/3 subgroup, IC1/2/3 subgroup, and all patients, respectively. The overall response rates to atezolizumab were 23\% (95\% CI: 15.6-31.9), 14.1\% (95\% CI: 10.4-18.5), and 13.4\% (95\% CI: 10.5-16.9) for the IC2/3 subgroup, IC1/2/3 subgroup, and all patients, respectively. These results indicate that PD-L1 status is a prognostic biomarker for both standard of care chemotherapies that comprised the control arm as well as atezolizumab in this patient population.

@@

```{r message = FALSE, fig.height = 6, fig.width = 8, include = knitr::is_latex_output(), eval = knitr::is_latex_output(), fig.cap = "Plot of decision rules made with ggplot. The color indicates whether the trial should stop or proceed for a given number of responses at each interim analysis."}
plot(two_sample_decision_tbl, plotly = FALSE)
```

```{r message = FALSE, fig.height = 6, fig.width = 8, include=knitr::is_html_output(), eval=knitr::is_html_output(), fig.cap = "Plot of decision rules made with plotly. The color indicates whether the trial should stop or proceed for a given number of responses at each interim analysis."}
plot(two_sample_decision_tbl)
```
